Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-31 numerical.tex
revising in progress by mkl

The modes shown in Fig. \ref{compare_modes_iso_kx10} are `body modes'
and are inertial waves destabilized by vertical shear
\citep{barker15}. (The possible exception being the fastest growing
mode for $\zmax=7H$ which is a surface mode --- see the discussion
below.)  The `fundamental corrugation mode', marked by `f',
corresponds to the eigenvalue with the smallest $|\sigma|$. It is robust
against vertical boundary conditions as it appears at the same
position in Fig. \ref{compare_modes_iso_kx10} for all cases. Higher  
order modes, corresponding to increasing $|\omega|$, increasingly
deviates from the analytic line due to the imposed boundaries, 
unless $\zmax$ is increased \citep{barker15}.

@@

In Fig. \ref{compare_modes_cool_kx30} we plot eigenvalues for unstable
modes with $\khat=30$. For clarity, we only plot analytic results 
for the fundamental mode ($M=0$) and $M=1,\,2$, and omit the
squares. The fundamental mode is again insensitive boundary
conditions compared to higher order modes, and are in close agreement
with analytic results. As before we observe more effective
stabilization for higher order modes, including the absence of surface
modes (which only occur for $\zmax\geq 5H$) as $\beta$ is increased.  
For $\zmax=3H$, the fundamental mode becomes dominant for $\beta\geq
0.03$. For $\zmax=5H$ and $\zmax=7H$, we
find modes slightly more unstable than the fundamental
mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
panels), but these are clearly associated with boundary
conditions as they are absent for $\zmax=3H$.

@@

Complications from vertical boundaries arise when considering large
$\khat$. For $\khat=30$ and $\zmax=5H,\,7H$ we find for $\beta\gtrsim
0.1$ the dominant disturbance correspond to the leftmost eigenvalues
seen in the lower two panels of Fig. \ref{compare_modes_cool_kx30},
which do not exist for $\zmax=3H$. For $\khat=50,\,100$ we find
surface modes dominate.

>>>>

The modes shown in Fig. \ref{compare_modes_iso_kx10} are `body modes'
and are inertial waves destabilized by vertical shear
\citep{barker15}. (The exception being the fastest growing
mode for $\zmax=7H$ which is a surface mode --- see the discussion
below.)  The `fundamental corrugation mode', marked by `f',
corresponds to the eigenvalue with the smallest $|\sigma|$. It is robust
against vertical boundary conditions as it appears at the same
position in Fig. \ref{compare_modes_iso_kx10} for all cases. Higher  
order modes, corresponding to increasing $|\omega|$, increasingly
deviates from the analytic line due to the imposed boundaries, 
unless $\zmax$ is increased \citep{barker15}.

@@

%presence of buoyancy forces at the boundary may make BC more
%influential

In Fig. \ref{compare_modes_cool_kx30} we plot eigenvalues for unstable 
modes with $\khat=30$. For clarity, we only plot analytic results 
for the fundamental mode ($M=0$) and $M=1,\,2$, and omit the 
squares. The fundamental mode is again insensitive to $\zmax$ and are
in close agreement with analytic results. However, 
there is increasing mismatch between analytical and numerical results
for higher order modes as $\beta$ is increased. This suggests that, in
the presence of buoyancy, boundary effects become important for these
modes, which have larger $\khat$ than that considered previously. For
example, for $\zmax=5H$ and $\zmax=7H$, we 
find modes slightly more unstable than the fundamental
mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
panels), but these are clearly affected by boundary
conditions as they are absent for $\zmax=3H$.

Nevertheless, as before we observe more effective stabilization for
higher order modes, including the absence of surface modes, for
example as $\beta$ is increased from $0.01$ to $0.03$ (for $\zmax\geq
5H$). If we discard modes associated with boundary conditions, then
the fundamental mode would again become the dominant mode as $\beta$ is
increased.

@@

Complications from vertical boundaries arise when considering large
$\khat$. For $\khat=30$ and $\zmax=5H,\,7H$ we find for $\beta\gtrsim
0.1$ the dominant disturbance correspond to the leftmost eigenvalues 
seen in the lower two panels of Fig. \ref{compare_modes_cool_kx30},
which do not exist for $\zmax=3H$. For $\khat=50,\,100$ and
$\beta\to0$, we find surface modes dominate so the maximum growth rate
depends on $\zmax$. However, the maximum growth rate is sensitive to
$\zmax$ for moderately large $\beta$ ($\gtrsim 0.1$), i.e. close to
the stabilization of the fundamental mode. 
%not surprising since we enable strongly stabilizing buoyancy forces
%at large zmax
